Exploring dependencies of multivariate seabird distributions: Linking graphical models to social network analysis

### 2017-07-03 paper1_jrss.tex
updated results

\usepackage{amssymb, amsmath, blindtext, enumitem}

\usepackage{longtable}

@@

\title[Graphical modeling of seabird distributions]{Graphical modeling of seabird distributions}
\author[Author 1 {\it et al.}]{Matthias Eckhardt}
\address{Affiliation,
City,
         Country.}
\email{Author@emailaddress.com}
\author{Earvin Balderama}
\address{Loyola University Chicago,
         Chicago, IL,
         USA.}

@@

@@

%============================================
\section{Methods}
%============================================
\begin{itemize}
\item We estimate the conditional spatial dependence structure between different seabird species
\item that is, structural dependence between two patterns conditional on all remaining patterns
\item so not interrelation between points $i$ and $j$, but interrelations between components $i$ and $j$ where $i$ and $j$ are sets of points within a bounded planar region.
\item edges express the partial (pairwise) interrelation between two species conditional on al remaining species
\item we omit to implement a formal test statistic as we assume that dependence might vary between species
\item we set a threshold indicating weak/ intermediate partial effects
\item we combine spatial dependence graph model with importance tools of social network analysis
\item info can provide important insides for species conservation
\item eg which are the most important species in the overall graph
\end{itemize}

@@

@@

>>>>

\usepackage{amssymb, amsmath, blindtext, enumitem}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{dsfont}

\usepackage{listings}

@@

\title[Graphical modeling of seabird distributions]{Exploring dependencies of multivariate seabird distributions: Linking graphical models to social network analysis}
\author[Author 1 {\it et al.}]{Matthias Eckardt}
\address{Humboldt-Universit{\"a}t zu Berlin,
Berlin,
 Germany.}
\email{Author@emailaddress.com}
\author{Earvin Balderama}
\address{Loyola University Chicago,
         Chicago, IL,
         USA.}

@@

\textit{ For our analysis we considered  on 128812 locations on marine bird sightings related to 306 different marine birds. Form this data, we preselected a sample of 204 species which have have at least been observed at three locations. By this, our analysis covered 128674 different locations.}

The observed planar point patterns for the $20$ most frequently recorded bird species is shown in Figure \ref{fig:birds20}.

\begin{figure}
\centering
\makebox{\includegraphics[scale=0.45]{BirdsTop20.eps}}
      \caption{\label{fig:birds20} Spatial point patterns of the $20$ most frequently observed marine bird species.}
\end{figure}

\textbf{Maybe reduce table to most frequent species, eg. top 20 birds}

@@

%============================================
\section{Methods}
%============================================

data as multivariate marked spatial point process as defined in \cite{Eckardt2016b}.

literature on point process
\cite{Chiu2013}, \cite{Illian2008} and \cite{StoyanStoyan1994}.
For a comprehensive treatment of various models and statistics for marks and points we refer the interested reader to \cite{Illian2008}, \cite{Baddeley2010} and \cite{WiegandMoloney2013}

\begin{itemize}
\item We estimate the conditional spatial dependence structure between different seabird species
\item that is, structural dependence between two patterns conditional on all remaining patterns
\item so not interrelation between points $i$ and $j$, but interrelations between components $i$ and $j$ where $i$ and $j$ are sets of points within a bounded planar region.
\item edges express the partial (pairwise) interrelation between two species conditional on al remaining species
\item we omit to implement a formal test statistic as we assume that dependence might vary between species
\item we set a threshold indicating weak/ intermediate partial effects
\item we combine spatial dependence graph model with importance tools of social network analysis
\item info can provide important insides for species conservation
\item eg which are the most important species in the overall graph
\end{itemize}

@@

For an intense description of different aspects of network analysis we refer the interested reader to \cite{Carrington2005}, \cite{Kolaczyk:2009} and \cite{Goldenberg:EtAl:2010}.

Although graphical modelling and network analysis are two distinct fields and are in general treated separately, the conjunction of both fields could offer additional benefits for the analysis of multivariate spatial data.  Especially for the analysis of high-dimensional spatial data one might be interested in the structural importance of a distinct node in a given SDGM. That is, how substantial is the impact of a certain node on the observed network structure. Queries of this type are covered by centrality measures which aim to value the importance of a distinct  node in the network. Although a rich methodological toolbox for the structural analysis of (social) network data exist, we only  we considered four different centrality measures: the degree, the closeness, the betweenness and the eigenvector centrality (cf. \cite{Borgatti2006, Freeman1979, Newman2004}). Although these measure are similar in spirit, they are different in detail and each of these three centrality measures formalizes a different concept of importance from a (social) network  perspective.

\textbf{meaning}
 Betweenness centrality measures the number of shortest paths
going through a specific vertex

Closeness is the mean geodesic distance between a given node and
 all other nodes with paths from the given node to the other
 node. This is close to being the mean shortest path, but 
geodesic distances give higher values for more central nodes.

 Eigenvector centrality gives greater weight to a node the more 
 it is connected to other highly connected nodes. A node
 connected to five high-scoring nodes will have higher 
 eigenvector centrality than a node connected to five low-scoring
 nodes. Thus, it is often interpreted as measuring a node's
 network

For our analysis we estimated all marked auto- and cross-spectral densities for all $204$ species nonparametrically using a discrete Fourier transform of the demeaned mark values. Form this, we computed separate spectra density matrices of dimension $204 \times 204$ for all frequencies. To obtain the mSDGM, we finally inverted and rescaled each matrix such that an edge in the mSDGM is equivalent to non-vanishing absolute rescaled marked inverse spectral value in at least one frequency. To capture variation in strength of spatial interdependencies in our data, we choose two arbitrary thresholds $\alpha$ for the edges below of which we assume two patterns as conditional uncorrelated. That is, as our data contains a wide range of differently-sized of marine bird species we assume a variation in strength of spatial interrelations between different types of trees. A similar idea has been proposed by \cite{GatherImhoffFried2002} with respect to intensive care monitoring.

computed two different mSDGMs by using two thresholds $\alpha=0.3$ and $\alpha=0.6$.

@@

For our first mSDGM, we considered a threshold $\alpha=0.3$ which we interpret as a conditional interrelation of weak effect size. For this, we treated all values of the empirical absolute rescaled marked inverse spectra below $\alpha=0.3$ as zero. The resulting mSDGM  is depicted in Figure \ref{fig:g30}.

\begin{figure}
\centering
\makebox{\includegraphics[scale=0.5]{rawg30.eps}}
      \caption{\label{fig:g30} Marked spatial dependence graph model for the $204$ different marine bird species, and bird counts as quantitative mark for a threshold level of $\alpha=0.3$.}
\end{figure}

This SDGM consisted of two $4$-node subgraphs, a $6$- node subgraph and a $153$-node subgraph and of 3400 edges. Besides, we also found $37$ isolated species (AMCO, AMPI, AMRE, BAOR, BLPW, BRAN, BTBW, BUPE, COYE, FEPE, GCKI, GRCA, LESC, LISH, MABO, MAWA, MERL, NOFL, NOHA NOTE, NOWA, PEFA, RBTR, RCKI, RTHU, RUBL, SAVS, SNBU, SPAR, Surf Scoter, UNHU, UNPT, UNTB, USAN, WTSP, YBCH, YWAR). These findings imply that non of the $37$ species are interrelated to any other species taken into account at threshold $\alpha=.3$. This could mean that the isolated nodes tend to represent species whose marked point patterns are different from all remaining marked point patterns taken into account. This is an interesting result as the isolated nodes are related to both frequent and endangered species and also differently-sized types of marine birds. One reason for this could be due to our sample scheme. We restricted the dataset to a subset of $204$ different marine bird species, so while there probably are interrelations of this $37$ aerial species to alternative types of birds, the alternative birds did not appear in our subsample. However, the global interrelations within coastal habitat are very complex and are still not completely understood.

Although several findings could be discussed in detail, we only considered structural interdependies for the marked patterns of Roseate Tern, Least Tern and the endangered species Black-capped Petrel, GBHE, YCNH and SAVS.

We begin our discussion with the structural interrelations for Roseate Tern and Least Tern. While we found $62$ conditionally interrelated marked patterns for Roseate Tern, we only observed $10$ interdependent species for Least Tern. For Roseate Tern, these species are Dovekie, AMGO, AMKE, AMOY, AMRO, BBPL, BHCO, BHGU, Red-necked Phalarope, Red Phalarope, BRBO, CHSW, DUNL, EUST, Forster's Tern, SBDO, GRYE, HADU, HOME, LEYE, NOPI, MALL,  Manx Shearwater, PAWA, PEEP, PESA, PISI, RBNU, REDH, RUDU, RUTU, SAGU, SATE, Unidentified Scaup, Pomarine Jaeger, Red-breasted Merganser, Ring-billed Gull, Great or Lesser Black-backed Gull, UCAT, UNBI, UNFR, UNPE, Unidentified Large Alcid , Unidentified Large Gull, Unidentified Merganser, Unidentified Small Tern, BANS, WHIM, GREG, MODO, Unidentified Storm-petrel, WIPH, WODU, YCNH, YRWA, BRSP, BLKI, Black-capped Petrel,   BonapArctic Tern's Gull, BRNO, Common Goldeneye and Dark Scoter. 
For Least Tern we have edges to   Common Murre, Barn Swallow, Arctic Tern, Unidentified Cormorant, Unidentified Shorebird, Unidentified Diving/Sea Duck, South Polar Skua, Unidentified Grebe, Black Guillemotand, Parasitic Jaeger.

For GBHE we observed $74$ direct interrelations to alternative marine birds, namely AMGO, AMKE, AMOY, AMRO, BBPL, BHCO, BHGU, GRCO, BRBO, CATE, CHSW, COME, DEJU, DUNL, EUST,  GADW, GRHE, SBDO, GRYE, HADU, HOME, LESA, LEYE, 
LIGU, NOPI, HEPE, ICGU, LTJA, MALL, OSPR,  PAWA, PEEP, PESA, PISI, PUMA, PUSA, RBNU, REDH,  RUDU, RUTU, RWBL, SAGU, SAND, SATE, SEPL,
 SESA, SOSP, SPSA, TRES, UCAT, ULSB, UNEI, UNFR, UNPE, BANS, UNWA, MYWA, WFSP, WHIM, GLGU,  GOME, GREG, MODO, BRTE, WILL, WIPH, WODU, YCNH, YRWA, UCRT, BLTE, BRNO, CANG, and, lastly, AWPE. \textbf{These birds are .....}.

Similarly, focussing on the interrelation which are present for YCNH, we found $79$ links. These interdependent species are  AWPE, Common Loon, Dovekie, Great Black-backed Gull, AMGO, AMKE, AMOY, AMRO, BBPL, BHCO, BHGU, Great Shearwater, Roseate Tern, BRBO, CHSW, COME, DEJU, DUNL, EUST, GADW, GRHE, GBHE, SBDO, GRYE,  HADU, HOME, LESA, LEYE, NOPI, ICGU, Laughing Gull, MALL, OSPR, PAWA, PEEP, PESA, PISI, PUSA, RBNU
 REDH, RUDU, RUTU, RWBL, SAGU, SATE, Unidentified Scaup, SEPL, SESA, SOSP, Pomarine Jaeger, Ring-billed Gull, SPSA, TRES, UCAT  ULSB, UNBI, UNEI, UNFR, UNPE, Unidentified Large Gull, Unidentified Merganser, Unidentified Small Tern, BANS, WFSP, WHIM, GLGU, GOME, GREG, MODO, WILL, WIPH, WODU, YRWA, UCRT, BLKI, Black-capped Petrel,   BonapArctic Tern's Gull, BRNO and CANG.\textbf{These birds are .....}.

For Black-capped Petrel, we found 64 alternative marked point patterns which are directly interrelated to the spatial marked pattern of Black-capped Petrel. These species are 
  Dovekie, AMGO, AMKE, AMOY, AMRO, BBPL, BHCO, BHGU, Red-necked Phalarope, Red Phalarope, Roseate Tern, BRBO, CHSW, DUNL, EUST, Forster's Tern, SBDO, GRYE, HADU, HOME, LEYE, NOPI, Horned Grebe, Laughing Gull, MALL,  Manx Shearwater, PAWA, PESA, PISI, RBNU,  REDH, RUDU, RUTU, SAGU, SATE, Unidentified Scaup, Pomarine Jaeger, Red-breasted Merganser,  Ring-billed Gull, Sooty Shearwater, Great or Lesser Black-backed Gull, UCAT, UNBI, UNFR, UNPE,
 Unidentified Large Alcid , Unidentified Large Gull, Unidentified Merganser, Unidentified Small Tern, BANS, WHIM, GREG, MODO,  Unidentified Storm-petrel, WIPH, WODU, YCNH, YRWA, BRSP, BLKI,
   BonapArctic Tern's Gull, BRNO, Common Goldeneye, and Dark Scoter. \textbf{These birds are .....}.

\textbf{Interestingly, SAVS is isolated in both SDGMs}

These findings imply the presence of complex conditional interdependencies between the marked spatial pattern of GBHF and also YCNH to a wide range of alternative species whereas we found no spatial interdependencies for SAVS to any alternative species taken into account.

As a second model, we also computed a SDGM for a threshold $\alpha=.6$. (we interprete this as interdependencies of intermediate effect size). The resulting mSDGM is depicted in Figure \ref{fig:g60}.

\begin{figure}
\centering
\makebox{\includegraphics[scale=0.5]{rawg60.eps}}
      \caption{\label{fig:g60} Marked spatial dependence graph model for the $204$ different marine bird species, and bird counts as quantitative mark for a threshold level of $\alpha=0.3$.}
\end{figure}

In this SGDM, we observed two pairs of joined nodes, a $3$-node subgraph, a $4$-node subgraph, a $6$-node subgraph, a $7$-node subgraph, a $8$-node subgraph, a $10$-node subgraph, a $49$-node subgraph and a , a $4$-node subgraph, a $6$-node subgraph, a $71$-node subgraph and $1880$ edges. In addition, we also found $42$ isolated nodes, namely AMCO, AMPI, AMRE, BAOR, BLPW, BRAN, BTBW, BUPE, COYE, FEPE, GCKI, GRCA, Herring Gull, LESC, LISH, MABO, MAFR, MAWA, MERL, NOFL, NOHA, NOTE, NOWA, PEFA, RBTR, RCKI, RTHU, RUBL, SAVS, SNBU, SPAR, Surf Scoter, SWAL, UNDU, UNHU, Unidentified Passerine, UNPT, UNTB, USAN, WTSP, YBCH, YWAR. Again, for the isolated nodes, we concluded that non of marked spatial patterns of these species is conditionally interdependent to any alternative marked pattern taken into account at a threshold $\alpha=.6$.

As before, we first discuss the spatial interdependencies for Roseate Tern and Least Tern. In contrast to our previous findings, only $13$ edges remain for Roseate Tern at a threshold $\alpha=.6$, namely Red-necked Phalarope, Forster's Tern, Unidentified Scaup, Pomarine Jaeger, Ring-billed Gull, Unidentified Merganser, Unidentified Small Tern, BRSP, BLKI, Black-capped Petrel,   BonapArctic Tern's Gull, Common Goldeneye, and Dark Scoter.

Least Tern is linked to 7 from 204 species, named:
   Common Murre Arctic Tern Unidentified Cormorant Unidentified Diving/Sea Duck South Polar Skua Black GuillemotParasitic Jaeger

\textbf{ .....}

With respect to the marked spatial pattern of GBHE, the neighborhood consisted of $41$ alternative species, namely AWPEBANS, BHCO, BHGU,  BLTE,  CATE, CANG, CHSW, COME, DEJU, DUNL, GLGU, GOME, GRCO, LESA, LIGU, ICGU, LTJA, MALL, MODO, MYWA, OSPR, PAWA, PEEP, PUMA, REDH, RUTU, RWBL, SAND, SEPL, SESA, SOSP, TRES, UNEI, UNFR, UNPE,  UNWA, YRWA, WFSP, WHIM, and WODU.

\textbf{...}

For YCNH we detected $52$ interrelated species at a threshold $\alpha=.6$. 
  AMGO, AMKE, AMOY, AMRO, BBPL, BHCO, BHGU, BRBO, CHSW, COME, DUNL, EUST, SBDO, GRYE, HADU, HOME, LESA, LEYE, NOPI, ICGU, MALL, OSPR, PAWA, PEEP, PESA, PISI, RBNU, REDH, RUDU, RUTU, RWBL, SAGU, SATE, SEPL, SESA, SOSP,
 UCAT, UNEI, UNFR, UNPE, BANS, WFSP, WHIM, GOME, GREG,
 MODO, WIPH, WODU, YRWA, BRNO, CANG and AWPE.

\textbf{...}

Similarly, only $13$ conditional interdependencies remained with respect to the marked pattern of  Black-capped Petrelfor a threshold $\alpha=.6$ including the marked point patterns of Red-necked Phalarope, Roseate Tern, Forster's Tern, Unidentified Scaup, Pomarine Jaeger, Ring-billed Gull, Unidentified Merganser, Unidentified Small Tern, BRSP, BLKI,   BonapArctic Tern's Gull, Common Goldeneye and also Dark Scoter.

To detect the most influential species within our data, we compute 4 different centrality measures (listed Table \ref{tab:centrality:measures}). The results for the threshold $\alpha=.3$ are listed below, isolated species excluded.

\begin{table}
\caption{\label{tab:centrality:measures}Centrality measures for SDGM for threshold $\alpha=.3$. Isolated nodes excluded.}
\centering
\fbox{%
\begin{tabular}{*{5}{c}}
\em code &\em degree &\em betweenness &\em closeness &\em eigen\\
MALL  &  $80$  &  $80.370354$  &  $9.37E-05$  &  $1.000000e+00$\\
BANS  &  $80$  &  $80.370354$  &  $9.37E-05$  &  $1.000000e+00$\\
WHIM  &  $80$  &  $80.370354$  &  $9.37E-05$  &  $1.000000e+00$\\
WODU  &  $80$  &  $80.370354$  &  $9.37E-05$  &  $1.000000e+00$\\
AMKE  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
AMRO  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
BBPL  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
BHCO  &  $79$  &  $74.852463$  &  $9.37E-05$  &  $9.917427e-01$\\
BRBO  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
CHSW  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
EUST  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
SBDO  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
GRYE  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
HADU  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
HOME  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
LEYE  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
NOPI  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
PAWA  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
PESA  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
PISI  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
RBNU  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
RUDU  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
RUTU  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
SAGU  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
SATE  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
UNFR  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
UNPE  &  $79$  &  $74.852463$  &  $9.37E-05$  &  $9.917427e-01$\\
GREG  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
WIPH  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
YCNH  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
BRNO  &  $79$  &  $64.521344$  &  $9.37E-05$  &  $9.930130e-01$\\
BHGU  &  $78$  &  $70.510241$  &  $9.37E-05$  &  $9.829660e-01$\\
DUNL  &  $78$  &  $70.510241$  &  $9.37E-05$  &  $9.829660e-01$\\
REDH  &  $78$  &  $70.510241$  &  $9.37E-05$  &  $9.829660e-01$\\
MODO  &  $78$  &  $70.510241$  &  $9.37E-05$  &  $9.829660e-01$\\
YRWA  &  $78$  &  $70.510241$  &  $9.37E-05$  &  $9.829660e-01$\\
TRES  &  $76$  &  $396.293914$  &  $9.32E-05$  &  $8.710167e-01$\\
SOSP  &  $75$  &  $45.678051$  &  $9.36E-05$  &  $9.546329e-01$\\
UNEI  &  $75$  &  $45.678051$  &  $9.36E-05$  &  $9.546329e-01$\\
DEJU  &  $74$  &  $191.068166$  &  $9.32E-05$  &  $8.695110e-01$\\
GBHE  &  $74$  &  $226.132882$  &  $9.32E-05$  &  $8.697880e-01$\\
GLGU  &  $74$  &  $191.068166$  &  $9.32E-05$  &  $8.695110e-01$\\
LESA  &  $73$  &  $125.782452$  &  $9.31E-05$  &  $8.690015e-01$\\
OSPR  &  $73$  &  $125.782452$  &  $9.31E-05$  &  $8.690015e-01$\\
AMGO  &  $72$  &  $49.578329$  &  $9.36E-05$  &  $9.162885e-01$\\
AMOY  &  $72$  &  $49.578329$  &  $9.36E-05$  &  $9.162885e-01$\\
COME  &  $72$  &  $104.948978$  &  $9.31E-05$  &  $8.675207e-01$\\
UCAT  &  $72$  &  $49.578329$  &  $9.36E-05$  &  $9.162885e-01$\\
WFSP  &  $72$  &  $104.948978$  &  $9.31E-05$  &  $8.675207e-01$\\
RWBL  &  $70$  &  $22.730622$  &  $9.34E-05$  &  $9.053099e-01$\\
SEPL  &  $70$  &  $22.730622$  &  $9.34E-05$  &  $9.053099e-01$\\
AWPE  &  $70$  &  $22.730622$  &  $9.34E-05$  &  $9.053099e-01$\\
PEEP  &  $69$  &  $38.941332$  &  $9.35E-05$  &  $8.863426e-01$\\
SESA  &  $69$  &  $56.090039$  &  $9.31E-05$  &  $8.620495e-01$\\
GOME  &  $69$  &  $56.090039$  &  $9.31E-05$  &  $8.620495e-01$\\
UNST  &  $66$  &  $864.535742$  &  $9.35E-05$  &  $7.261004e-01$\\
UNBI  &  $65$  &  $162.419996$  &  $9.35E-05$  &  $7.316740e-01$\\
LAGU  &  $64$  &  $388.758703$  &  $9.35E-05$  &  $6.854566e-01$\\
POJA  &  $64$  &  $157.699639$  &  $9.34E-05$  &  $7.174345e-01$\\
BCPE  &  $64$  &  $979.935364$  &  $9.36E-05$  &  $6.892403e-01$\\
CANG  &  $64$  &  $23.269845$  &  $9.35E-05$  &  $8.382226e-01$\\
GBBG  &  $63$  &  $511.095976$  &  $9.35E-05$  &  $6.644262e-01$\\
SCAU  &  $63$  &  $429.075802$  &  $9.35E-05$  &  $7.079279e-01$\\
RBGU  &  $63$  &  $174.4483$  &  $9.35E-05$  &  $7.533343e-01$\\
UNME  &  $63$  &  $153.685067$  &  $9.34E-05$  &  $7.091772e-01$\\
BOGU  &  $63$  &  $174.4483$  &  $9.35E-05$  &  $7.533343e-01$\\
ROST  &  $62$  &  $148.609153$  &  $9.34E-05$  &  $6.905912e-01$\\
BLKI  &  $62$  &  $165.552583$  &  $9.34E-05$  &  $7.510736e-01$\\
COLO  &  $60$  &  $454.304315$  &  $9.34E-05$  &  $6.354029e-01$\\
GADW  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
GRHE  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
PUSA  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
SPSA  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
ULSB  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
WILL  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
UCRT  &  $59$  &  $1.436417$  &  $9.30E-05$  &  $7.795919e-01$\\
ICGU  &  $58$  &  $1.649017$  &  $9.30E-05$  &  $7.766811e-01$\\
UNLG  &  $58$  &  $157.442443$  &  $9.34E-05$  &  $6.242535e-01$\\
DOVE  &  $56$  &  $194.321598$  &  $9.34E-05$  &  $5.873140e-01$\\
GRSH  &  $48$  &  $288.105101$  &  $9.33E-05$  &  $4.697169e-01$\\
LIGU  &  $48$  &  $123.565465$  &  $9.29E-05$  &  $4.833385e-01$\\
MYWA  &  $48$  &  $123.565465$  &  $9.29E-05$  &  $4.833385e-01$\\
BRSP  &  $28$  &  $407.588982$  &  $9.30E-05$  &  $1.731580e-01$\\
RNPH  &  $27$  &  $53.421704$  &  $9.29E-05$  &  $1.731189e-01$\\
UNSP  &  $27$  &  $49.85151$  &  $9.29E-05$  &  $1.549736e-01$\\
COGO  &  $26$  &  $39.864962$  &  $9.29E-05$  &  $1.722011e-01$\\
CATE  &  $24$  &  $25.047311$  &  $9.21E-05$  &  $1.539562e-01$\\
FOTE  &  $24$  &  $24.555419$  &  $9.29E-05$  &  $1.607928e-01$\\
BLTE  &  $24$  &  $25.047311$  &  $9.21E-05$  &  $1.539562e-01$\\
LTJA  &  $23$  &  $22.428263$  &  $9.21E-05$  &  $1.534467e-01$\\
UNWA  &  $23$  &  $22.428263$  &  $9.21E-05$  &  $1.534467e-01$\\
DASC  &  $23$  &  $246.349381$  &  $9.29E-05$  &  $1.550616e-01$\\
GRCO  &  $22$  &  $19.270691$  &  $9.21E-05$  &  $1.297163e-01$\\
PUMA  &  $22$  &  $19.270691$  &  $9.21E-05$  &  $1.297163e-01$\\
SAND  &  $22$  &  $19.270691$  &  $9.21E-05$  &  $1.297163e-01$\\
SOSH  &  $22$  &  $35.136566$  &  $9.28E-05$  &  $9.390908e-02$\\
UNDD  &  $21$  &  $431.447758$  &  $9.29E-05$  &  $6.035436e-02$\\
UNLT  &  $21$  &  $20.900229$  &  $9.28E-05$  &  $8.421874e-02$\\
BRTE  &  $20$  &  $14.69234$  &  $9.21E-05$  &  $1.053226e-01$\\
UNCO  &  $19$  &  $410.715654$  &  $9.28E-05$  &  $6.528103e-02$\\
BLSC  &  $18$  &  $428.995063$  &  $9.28E-05$  &  $7.086913e-02$\\
DCCO  &  $18$  &  $245.042857$  &  $9.27E-05$  &  $6.999939e-02$\\
LTDU  &  $17$  &  $41.640397$  &  $9.27E-05$  &  $7.254790e-02$\\
NOGA  &  $17$  &  $160.237757$  &  $9.27E-05$  &  $5.381070e-02$\\
HEPE  &  $16$  &  $155.745163$  &  $9.20E-05$  &  $5.593917e-02$\\
MASH  &  $16$  &  $90.089212$  &  $9.27E-05$  &  $6.768689e-02$\\
REPH  &  $15$  &  $59.456976$  &  $9.27E-05$  &  $6.550681e-02$\\
RBME  &  $15$  &  $59.456976$  &  $9.27E-05$  &  $6.550681e-02$\\
UBBG  &  $15$  &  $59.456976$  &  $9.27E-05$  &  $6.550681e-02$\\
UNLA  &  $15$  &  $59.456976$  &  $9.27E-05$  &  $6.550681e-02$\\
ABDU  &  $14$  &  $2.912409$  &  $9.20E-05$  &  $4.369921e-02$\\
WTTR  &  $14$  &  $2.912409$  &  $9.20E-05$  &  $4.369921e-02$\\
COMU  &  $13$  &  $133.71799$  &  $9.20E-05$  &  $3.216736e-03$\\
RAZO  &  $13$  &  $368.994253$  &  $9.25E-05$  &  $3.164179e-02$\\
BARS  &  $12$  &  $207.723894$  &  $9.17E-05$  &  $3.455434e-04$\\
SPSK  &  $12$  &  $53.468537$  &  $9.20E-05$  &  $4.227004e-03$\\
UNGR  &  $12$  &  $207.723894$  &  $9.17E-05$  &  $3.455434e-04$\\
UNSG  &  $12$  &  $224.536021$  &  $9.24E-05$  &  $1.560035e-02$\\
BRPE  &  $12$  &  $224.536021$  &  $9.24E-05$  &  $1.560035e-02$\\
RNGR  &  $11$  &  $0$  &  $9.09E-05$  &  $1.786223e-02$\\
SHOR  &  $11$  &  $174.360858$  &  $9.16E-05$  &  $3.003178e-04$\\
ATPU  &  $11$  &  $39.152444$  &  $9.16E-05$  &  $5.391765e-03$\\
ARTE  &  $10$  &  $39.727535$  &  $9.19E-05$  &  $1.996873e-03$\\
GRSK  &  $10$  &  $268.176822$  &  $9.20E-05$  &  $2.782868e-03$\\
LESP  &  $10$  &  $310.367803$  &  $9.25E-05$  &  $3.691295e-02$\\
LETE  &  $10$  &  $39.727535$  &  $9.19E-05$  &  $1.996873e-03$\\
HOGR  &  $9$  &  $720.079655$  &  $9.27E-05$  &  $1.088519e-02$\\
SOTE  &  $9$  &  $132.848369$  &  $9.20E-05$  &  $2.527661e-03$\\
AUSH  &  $9$  &  $3.172553$  &  $9.23E-05$  &  $1.557767e-02$\\
LBBG  &  $8$  &  $21.656632$  &  $9.09E-05$  &  $9.112791e-05$\\
MAFR  &  $8$  &  $0$  &  $9.20E-05$  &  $3.624032e-02$\\
TBMU  &  $8$  &  $149$  &  $9.05E-05$  &  $1.647815e-05$\\
UNSK  &  $8$  &  $149$  &  $9.05E-05$  &  $1.647815e-05$\\
BLGU  &  $8$  &  $151$  &  $9.18E-05$  &  $1.983332e-03$\\
COSH  &  $7$  &  $89.976832$  &  $9.17E-05$  &  $1.506881e-03$\\
UNJA  &  $7$  &  $10.767651$  &  $9.09E-05$  &  $5.559037e-05$\\
UNTE  &  $7$  &  $89.976832$  &  $9.17E-05$  &  $1.506881e-03$\\
PAJA  &  $7$  &  $0$  &  $9.18E-05$  &  $1.982934e-03$\\
ROYT  &  $6$  &  $23.486786$  &  $9.12E-05$  &  $5.376720e-04$\\
WWSC  &  $6$  &  $23.564037$  &  $9.15E-05$  &  $5.104993e-04$\\
UNPH  &  $6$  &  $23.486786$  &  $9.12E-05$  &  $5.376720e-04$\\
UNSH  &  $6$  &  $23.486786$  &  $9.12E-05$  &  $5.376720e-04$\\
CEDW  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
GBTE  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
SNGO  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
UNEG  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
UNSA  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
UNFI  &  $5$  &  $0$  &  $2.48E-05$  &  $0.000000e+00$\\
COTE  &  $5$  &  $74.509944$  &  $9.15E-05$  &  $5.842575e-04$\\
WISP  &  $5$  &  $12.356908$  &  $9.13E-05$  &  $5.022849e-04$\\
RTLO  &  $4$  &  $0$  &  $9.05E-05$  &  $5.741835e-05$\\
UNAL  &  $4$  &  $25.213789$  &  $9.05E-05$  &  $3.724320e-05$\\
UNSC  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
GRSC  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
GWTE  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
UNGO  &  $3$  &  $0$  &  $8.93E-05$  &  $4.767535e-07$\\
UNDU  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
UNGU  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
UNLO  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
UNMU  &  $3$  &  $0$  &  $8.93E-05$  &  $4.767535e-07$\\
USSB  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
COEI  &  $3$  &  $0$  &  $2.45E-05$  &  $0.000000e+00$\\
NOFU  &  $2$  &  $0$  &  $9.13E-05$  &  $5.347083e-04$\\
BUFF  &  $2$  &  $99.767649$  &  $9.15E-05$  &  $9.987188e-04$\\
UNPA  &  $1$  &  $0$  &  $9.05E-05$  &  $2.828221e-05$\\
SWAL  &  $1$  &  $0$  &  $9.08E-05$  &  $7.976898e-04$\\
HERG  &  $1$  &  $0$  &  $9.15E-05$  &  $7.673378e-04$\\
\end{tabular}}
\end{table}

betweenness centrality is shown in Figure \ref{fig:g30between}

\begin{figure}
\centering
\makebox{\includegraphics[scale=0.5]{betweeng30.eps}}
      \caption{\label{fig:g30between} Betweenness centrality for the marked spatial dependence graph model for the $204$ different marine bird species, and bird counts as quantitative mark for a threshold level of $\alpha=0.3$.  
 }
\end{figure}

For this plot, we found that BCPE is the most important marine bird species by means of the betweenness centrality measure. Besides BCPE, we observed a high importance of UNST, HOGR as well as GBBG amongst other species.

Similar calculations have been performed for the threshold $\alpha=0.6$.

 The results are given below, isolated species excluded.

\begin{lstlisting}
39     COME     58   49.987456 3.674174e-05 1.000000e+00
59     LESA     58   49.987456 3.674174e-05 1.000000e+00
82     OSPR     58   49.987456 3.674174e-05 1.000000e+00
104    SESA     58   49.987456 3.674174e-05 1.000000e+00
155    WFSP     58   49.987456 3.674174e-05 1.000000e+00
160    GOME     58   49.987456 3.674174e-05 1.000000e+00
